---
title: "chebInterp Introduction"
author: "Walter Zhang"
date: "February 8, 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{chebInterp Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(chebInterp)
options(knitr.kable.NA = '')
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

In this vignette, we demonstrate how to use the chebInterp package to interpolate functions. The original use case of these functions was to interpolate the value function for a dynamic discrete choice problem. However, we will show how to us the functions in the package to interpolate simple functions. Let us consider the following three functions $f,g,h:\mathbb{R}^3 \mapsto \mathbb{R}$ that are defined as follows:
$$f(x_1,x_2,x_3) = x_1x_3^3 + x_2x_3 + x_1^2x_2x_3^2 $$ 
$$g(x_1, x_2, x_3) = x_1 \log(5+x_2)x_3 $$
$$h(x_1,x_2,x_3) = x_1^2\cos(x_2)\exp(x_3)$$
We also define $\mathbf{x} = [x_1,x_2,x_3]$. We want to approximate each function on the rectangle $([-5,2],[-2,4],[-3,3])$. To gauge approximation accuracy, we will then compare the Chebyshev approximation to a Monte Carlo evaluation at 10,000 uniforms draws on the bounded rectangle. $M$ is defined as interpolation nodes the and $N$ is number of degrees for the approximation.

```{r}
main   <-   function(args)
{
    bounds    <-   args$bounds
    upper_b   <-   unlist(lapply(bounds, function(ii) ii[2]))
    lower_b   <-   unlist(lapply(bounds, function(ii) ii[1]))
    D   <-  args$D
    N   <-  args$N
    M   <-  args$M
    
    set.seed(args$base_seed)
    # Simulate Draws
    x_draws   <-   sapply(1:length(upper_b), function(i)
    {
        runif(args$draws, min = lower_b[i], max = upper_b[i])
    })
    
    # Initialize the Chebyshev Approximation
    cheb_init   <-   initializeChebyshevApproximator(D, N, M, upper_b = upper_b, lower_b = lower_b)
    
    # Compute Chebyshev Approximation Coefficients
    cheb_function   <-   lapply(args$functions, function(func)
    {
        calculateChebyshevCoefficients(func, cheb_init, tolerance = args$tolerance)
    })
    
    # Evaluate the Chebyshev Approximation
    cheb_values   <-   lapply(cheb_function, function(func_coef)
    {
        evaluateChebyshev(x_draws, func_coef)
    })
    
    # Evaluate the Functions
    x_draws_df        <-   data.frame(x_draws)
    function_values   <-   lapply(args$functions, function(func) func(x_draws_df)[,1])
    
    # Compute Summary of Absolute Deviations
    summary_stats   <-   do.call(rbind, lapply(names(args$functions), function(func_name)
    {
        abs_deviation   <-   abs(cheb_values[[func_name]] - function_values[[func_name]])
        data.frame(`Function`   =   func_name,
                   Mean         =   mean(abs_deviation),
                   SD           =   sd(abs_deviation),
                   Max          =   max(abs_deviation),
                   D   =   args$D,
                   N   =   args$N,
                   M   =   args$M
                   )
    }))
    colnames(summary_stats)[1]   <-   "Function"

    return(summary_stats)
}
```


```{r}
args <- list()
args$base_seed <- 1234L
args$bounds    <-   list(c(-5, 2), c(-2, 4), c(-3, 3))
args$draws     <-   10000L

args$D           <-   length(args$bounds)
args$N           <- 3L
args$M           <- 7L
args$tolerance   <-   1e-12

args$functions   <-   list()
args$functions[["f"]]   <-   function(x) x[1]*x[3]^3 + x[2]*x[3] + x[1]^2*x[2]*x[3]^2
args$functions[["g"]]   <-   function(x) x[1]*log(5 + x[2])*x[3]
args$functions[["h"]]   <-   function(x) x[1]^2*cos(x[2])*exp(x[3])
```

With $N=3$ and $M=3$, we can compute the approximation error of the Chebyshev Interpolation for our functions. The Mean, SD, and Max are the respective mean. standard deviation. and maximum errors across the $10,000$ draws.

```{r}
knitr::kable(main(args), digits = 3)
```

We now examine the approximation error of function $g$ for a grid of $M$ and $N$ values. Specifically, we will consider $M\in\{1,\ldots,5\}$ and $N\in\{2,\ldots,7\}$.

```{r}
args$functions <- args$functions["g"]
parameter_grid   <-   expand.grid(1:5L, 2:7L)
colnames(parameter_grid)   <-   c("N", "M")
parameter_grid   <-   subset(parameter_grid, M > N)
```

```{r}
error_list   <-   lapply(1:nrow(parameter_grid), function(i)
{
    args_batch           <-   args
    args_batch$N         <-   parameter_grid[i, "N"]
    args_batch$M         <-   parameter_grid[i, "M"]

    main(args_batch)
})
error_DF <- do.call(rbind, error_list)
```

```{r results = 'asis'}
knitr::kable(reshape::cast(error_DF[c("N", "M", "Mean")], N~M, value = "Mean"), 
             digits = 3,
             caption = "Mean Approximation Error (N by M)")
```
